\vspace{-0.5in}{\bfseries\scshape Transfer Learning for computationally limited training of Vision Transformers on art classification tasks } \\ \vspace{1ex} \normalsize{Bachelor's Project Thesis}\vspace{-2ex}

### 2022-08-09 writing/tex/conclusion.tex
Incorporated feedback from Dr. Sabatelli. Now its time to REALLY submit itgit add -A! :-D

\section{Conclusion}
Seeing which is better, VTs or CNNs, has never been the goal of this paper. The underlying architecture is somewhat irrelevant, as one should always pick (or create) the model that fits the task best -- regardless of whether it is a VT, CNN, or something else. What this paper, then, was interested in, is whether VTs can become a valuable alternative to CNNs under circumstances in which:
\begin{itemize}
\item small datasets are used: specifically, small enough to have been hand-labeled by one person;
\item only consumer-grade hardware is available.
\end{itemize}
In other words: this paper was interested in the type of machine learning one could try at home, and wanted to know if VTs could be made useful for it.

>>>>

\section{Conclusion}
Seeing which is better, VTs or CNNs, has never been the goal of this paper. The underlying architecture is somewhat irrelevant, as one should always pick (or create) the model that fits the task best -- regardless of whether it is a VT, or CNN. What this paper, then, was interested in, is whether VTs can become a valuable alternative to CNNs under circumstances in which:
\begin{itemize}
\item small datasets are used: specifically, small enough to have been hand-labeled by one person;
\item only consumer-grade hardware is available.
\end{itemize}
In other words: this paper was interested in the type of machine learning one could try at home, and wanted to know if VTs could be made useful for it.

### 2022-08-09 writing/tex/experiments.tex
Incorporated feedback from Dr. Sabatelli. Now its time to REALLY submit itgit add -A! :-D

\section{Experiments}
This section focuses on the results gotten from the conducted experiments. These will first briefly be presented in section \ref{results}. After that, section \ref{exp:int} will discuss them in more detail, and see how they compare to earlier studies, etc.

@@

Supporting `evidence' might be visible in Figure \ref{results:img:sal}. It plots saliency maps for one CNN (\texttt{ConvNext}), and two VTs (\texttt{DeiT} and \texttt{Swin}). It is interesting that \texttt{ConvNext} shows much bigger differences between the OTS and FT plots. In Figure \ref{results:img:sal:picture}, for example, it learns to just look at the picture frame after FT. In Figure \ref{results:img:sal:dish} it learns to look more at the center, which allows it to correctly classify the input as a dish instead of a plate\footnote{No idea what the difference is between these classes, but these models do seem to know, so that is good!}. % TODO Maybe delete this comment, not very professional
These bigger differences for \texttt{ConvNext} align well with observations made earlier, as they suggest that there is still much improvement possible after just OTS learning for CNNs.

>>>>

\section{Experiments}
This section focuses on the results obtained from the conducted experiments. These will first briefly be presented in section \ref{results}. After that, section \ref{exp:int} will discuss them in more detail, and see how they compare to earlier studies.

@@

Supporting `evidence' might be visible in Figure \ref{results:img:sal}. It plots saliency maps for one CNN (\texttt{ConvNext}), and two VTs (\texttt{DeiT} and \texttt{Swin}). It is interesting that \texttt{ConvNext} shows much bigger differences between the OTS and FT plots. In Figure \ref{results:img:sal:picture}, for example, it learns to just look at the picture frame after FT. In Figure \ref{results:img:sal:dish} it learns to look more at the center, which allows it to correctly classify the input as a dish instead of a plate. These bigger differences for \texttt{ConvNext} align well with observations made earlier, as they suggest that there is still much improvement possible after just OTS learning for CNNs.

### 2022-08-09 writing/tex/introduction.tex
Incorporated feedback from Dr. Sabatelli. Now its time to REALLY submit itgit add -A! :-D

CNNs are largely invariant to shifts and small distortions of the input image. More important is that they can be seen as automatic feature extractors. It is well established (e.g. \citeauthor{lecun1998gradient}, \citeyear{lecun1998gradient}; \citeauthor{yosinski2014transferable}, \citeyear{yosinski2014transferable}) that the first layers of a CNN detect low-level features, such as edges and corners. These are combined in subsequent layers to detect higher-order features. The final layer should then give a representative summary of the input, such that, for example, an MLP-type layer can extract all the information it needs from it to perform the task at hand.

@@

The self-attention mechanism plays a central role in VTs, and is worth elaborating on. Each head of a self-attention layer contains 3 learned projection matrices, which are used to produce $key$, $value$, and $query$ vectors for every position in the input sequence. The output for a single position is a weighted sum of all $value$ vectors. Weights are obtained by matching this position's $query$ with all $key$ vectors\footnote{In principle by taking softmax over all query-key inner products.}. As such, an attention head decides itself how much of each position should be incorporated in subsequent representations.

The original Transformer architecture was designed to solve problems in \textit{natural language processing} (a field it now dominates). To make it work for CV applications, VTs divide an image into square patches, which get treated as tokens (words). Learned embeddings convert these into a sequence of vectors, to which 1D positional embeddings are added\footnote{Do mind that VTs have no built-in knowledge of an image's topology, as the architecture ignores the 2D layout.}. A learned $[class]$ vector gets prepended to the sequence, and its representation at the end of the encoder stack is the final output, which can then be used for classification, segmentation, etc.

@@

\paragraph{CNNs versus VTs}
Where CNNs reduce the number of trainable weights by taking an image's topology into account, VTs do this by learning themselves what information they should incorporate in successive representations of the input. As such, they are not restricted to a local receptive field, but can potentially include information from all over the previous layer.

>>>>

CNNs are largely invariant to shifts and small distortions of the input image. More important is that they can be seen as automatic feature extractors. It is well established (e.g. \citeauthor{lecun1998gradient}, \citeyear{lecun1998gradient}; \citeauthor{yosinski2014transferable}, \citeyear{yosinski2014transferable}) that the first layers of a CNN detect low-level features, such as edges and corners. These are combined in subsequent layers to detect higher-order features. The final layer should then give a representative summary of the input, such that, for example, a Multilayer Perceptron-type (MLP-type) layer can extract all the information it needs from it to perform the task at hand.

@@

The self-attention mechanism plays a central role in VTs, and is worth elaborating on. Each head of a self-attention layer contains three learned projection matrices, which are used to produce $key$, $value$, and $query$ vectors for every position in the input sequence. The output for a single position is a weighted sum of all $value$ vectors. Weights are obtained by matching this position's $query$ with all $key$ vectors\footnote{In principle by taking the softmax over all query-key inner products.}. As such, an attention head decides itself how much of each position should be incorporated in subsequent representations.

The original Transformer architecture was designed to solve problems in \textit{natural language processing} -- a field it now dominates \citep{wolf2020transformers}. To make it work for CV applications, VTs divide an image into square patches, which get treated as tokens (words). Learned embeddings convert these into a sequence of vectors, to which 1D positional embeddings are added\footnote{Do mind that VTs have no built-in knowledge of an image's topology, as the architecture ignores the 2D layout.}. A learned $[class]$ vector gets prepended to the sequence, and its representation at the end of the encoder stack is the final output, which can then be used for classification, segmentation, etc.

@@

\paragraph{CNNs versus VTs}
While CNNs reduce the number of trainable weights by taking an image's topology into account, VTs do this by learning themselves what information they should incorporate in successive representations of the input. As such, they are not restricted to a local receptive field, but can potentially include information from all over the previous layer.

### 2022-08-09 writing/tex/methods.tex
Incorporated feedback from Dr. Sabatelli. Now its time to REALLY submit itgit add -A! :-D

\subsection{Models} \label{methods:models}
Eight different neural network architectures are used throughout this paper, of which four are CNN-based and four VT-based. When multiple versions are available (i.e. `base', `small', `tiny'), `base'/`medium' ones are chosen.

>>>>

\subsection{Models} \label{methods:models}
Eight different neural network architectures are used, of which four are CNN-based and four VT-based. When multiple versions are available (i.e. `base', `small', `tiny'), `base'/`medium' ones are chosen.
